\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage{amsmath, amsthm}
\begin{document}
\author{Benjamin Sorenson} \title{Assignment 4}
\maketitle
\begin{enumerate}[label=\bfseries Question \arabic*:]
\item
  \begin{enumerate}
  \item Sentences in CNF (for sentence number two, I took it to mean
    ``It is not hot, and if it's sunny, then it's pleasant'' instead
    of ``If it's hot and sunny, then it's pleasnt'':
    \begin{enumerate}[label= KB\arabic*.]
    \item \(\lnot Humid \lor \lnot Hot \lor Sticky \)
    \item \(\lnot Hot\)
    \item \(\lnot Sunny \lor Pleasant\)
    \item \(Sunny\)
    \item \(Humid\)
    \item\(Sunny \lor Pleasant\)
    \end{enumerate}
  \item
    \begin{enumerate}
    \item \(\lnot Sticky\) (assumption)
    \item \(\lnot Pleasant\) (assumption)
    \item \(Pleasant \) (from KB3 and KB4)
    \end{enumerate}
  \item
    \begin{enumerate}
    \item \(\lnot Sticky \lor \lnot Pleasant\) (assumption)
    \item \(Pleasant \) (from KB3 and KB4)
    \item \(\lnot Sticky\) (from ii. and i.)
    \item \dots Here is where I run into a dead end. I don't think
      this can be proven from the collection of sentences. We can
      conclude that it's either \(Sticky\)
      or \(Pleasant\),
      but we can't say that it is both \(Sticky\)
      and \(Pleasant\).
      It could be that it's both \(Sticky\)
      and \(Pleasant\),
      but we can't tell from this collection of sentences.  In other
      words, the statement \(Sticky \land Pleasant\)
      is satisfiable, but not valid in this context.
    \end{enumerate}
  \end{enumerate}
\item
  \begin{proof} Let \(S\)
    be a statement in CNF such that
    \(S = P_1 \lor P_2 \lor \dots \lor P_n \lor Q_1 \lor Q_2 \lor
    \dots \lor Q_m \) then
    \begin{align*}
      S = (P_1 \lor P_2 \lor \dots \lor P_n) \lor (Q_1 \lor Q_2 \lor
      \dots \lor Q_m)       
      &= \lnot (P_1 \lor P_2 \lor \dots \lor P_n) \rightarrow (Q_1 \lor Q_2 \lor
        \dots \lor Q_m)                    
      \\ &= (\lnot P_1 \land \lnot P_2 \land \dots \land \lnot P_n) \rightarrow (Q_1 \lor Q_2 \lor
           \dots \lor Q_m) 
    \end{align*}
  \end{proof}
\item
  \begin{enumerate}
  \item
    \(\forall x ((Big(x) \land House(x)) \rightarrow Expensive(x))\)
  \item
    \(\forall x ((Expensive(x) \land House(x)) \rightarrow Big(x))\)
  \item
    \(\forall x ((Big(x) \land House(x)) \rightarrow Expensive(x))\)
  \item
    \(\exists x \exists y (Big(x) \land House(x) \land Condo(y) \land
    (Price(y) > Price(x))) \)
  \item
    \((\exists x \exists y (House(x) \land Garden(y) \land Has(x, y)))
    \land (\forall x \forall y \forall m \forall n (House(x) \land
    Garden(y) \land Has(x, y) \land House(m) \land Garden(n) \land
    Has(m, n)) \rightarrow x = m) \)
  \end{enumerate}
\item
  \begin{enumerate}
  \item Incorrect.  As written, the translation says that there is at
    least one pink house in Minneapolis, but does not exclude the
    possibility that there could be others.  It should be
    \(\exists x (house(x) \land in(x, Minneapolis) \land color(x,
    Pink)) \land \forall x \forall y (house(x) \land in(x,
    Minneapolis) \land color(x, Pink) \land house(y) \land in(y,
    Minneapolis) \land color(y, Pink)) \rightarrow x = y\)
  \item Incorrect. As written, the translation says that for every
    apartment there is some house that is bigger.  It should be
    \(\exists y \forall x( house(y) \land (appartment(x) \rightarrow
    bigger(y, x)))\)
  \item Incorrect.  As written, the translation says that for every
    apartment, there is some house that is cheaper. It should be
    \(\forall x \forall y (house(x) \land apartment(y)) \rightarrow
    cheaper(x, y)\)
  \item Incorrect.  As written, the translation says that there is
    some farm and some object such that if the object is a house then
    the farm is cheaper than it. It should be
    \(\exists x \exists y (farm(x) \land house(y) \land cheaper(x,
    y))\)
  \item Incorrect.  As written, the translation says that for every
    house, if there is a bathroom, then it is in the house.  It should
    be \(\forall x \exists y (house(x) bathroom(y) \land in(x, y))\)
  \end{enumerate}
\item Proof by resolution:
  \begin{enumerate}[label=\arabic*.]
  \item \(G(B)\)
  \item \(\lnot G(x) \lor H(x)\)
  \item \(\lnot H(z) \lor I(z)\)
  \item \(\lnot H(w) \lor J(w, D)\)
  \item \(\lnot I(B) \lor J(C, B)\)
  \item \(\lnot I(q) \lor \lnot J(q, y)\)
  \item \(\lnot G(B) \lor H(B)\) (from 2, \(B/x\))
  \item \(H(B)\) (from 1 and 7)
  \item \(\lnot H(B) \lor J(B, D)\) (from 4, \(B/w\))
  \item \(\lnot H(B) \lor I(B)\) (from 3, \(B/w\))
  \item \(I(B)\) (from 10 and 8)
  \item \(J(B, D)\) (from 8 and 9)
  \item \(\lnot I(B) \lor \lnot J(B, y)\) (from 6, \(B/q\))
  \item \(\lnot I(B) \lor \lnot J(B, D)\) (from 13, \(D/y\))
  \item \(\lnot J(B, D)\) (from 11 and 14)
  \item \(J(B, D)\land \lnot J(B, D)\) (from 12 and 15)
  \end{enumerate}
\item
  \begin{enumerate}
  \item Statements in predicate calculus:
    \begin{enumerate}[label=\arabic*.]
    \item\(\forall x (Person(x) \land Rich(x) \land \lnot Stupid(x))
      \rightarrow Happy(x)\)
    \item
      \( \forall x (Person(x) \land Reads(x)) \rightarrow \lnot
      Stupid(x) \)
    \item \(Person(JOHN) \land Reads(JOHN) \land Rich(JOHN) \)
    \item
      \(\forall x (Happy(x) \land Person(x)) \rightarrow
      HasExcitingLife(x) \)
    \end{enumerate}
  \item Conjunctive normal form:
    \begin{enumerate}[label=KB\arabic*.]
    \item
      \(\lnot Person(x) \lor \lnot Rich(x) \lor Stupid(x) \lor
      Happy(x)\)
    \item \(\lnot Person(y) \lor \lnot Reads(y) \lor \lnot Stupid(y)\)
    \item \(Person(JOHN)\)
    \item \(Reads(JOHN)\)
    \item \(Rich(JOHN)\)
    \item
      \(\lnot Happy(z) \lor \lnot Person(z) \lor HasExcitingLife(z)\)
    \end{enumerate}
\item Proof of \(\exists m HasExcitingLife(m)\):
  \begin{enumerate}[label=\arabic*.]
  \item \( \lnot HasExcitingLife(m)\) (assumption)
  \item
    \(\lnot Person(JOHN) \lor \lnot Rich(JOHN) \lor Stupid(JOHN) \lor
    Happy(JOHN)\) (from KB1, \(JOHN / x\))
  \item
    \(\lnot Person(JOHN) \lor \lnot Reads(JOHN) \lor \lnot
    Stupid(JOHN)\) (from KB2, \(JOHN / y\))
  \item
    \(\lnot Happy(JOHN) \lor \lnot Person(JOHN) \lor
    HasExcitingLife(JOHN)\) (from KB6, \(JOHN /z\))
  \item \( \lnot HasExcitingLife(JOHN)\) (from 1, \(JOHN / m\))
  \item \(\lnot Rich(JOHN) \lor Stupid(JOHN) \lor Happy(JOHN)\)
    (from KB3 and 1)
  \item \(Stupid(JOHN) \lor Happy(JOHN)\) (from 6 and KB5)
  \item \(\lnot Reads(JOHN) \lor \lnot Stupid(JOHN)\) (from 3 and KB3)
  \item \(\lnot Happy(JOHN) \lor HasExcitingLife(JOHN)\)
    (from 4 and KB3)
  \item \(\lnot Stupid(JOHN)\) (from KB4 and 8)
  \item \(Happy(JOHN)\) (from 7 and 10)
  \item \(HasExcitingLife(JOHN)\) (from 9 and 11)
  \item \(HasExcitingLife(JOHN) \land \lnot HasExcitingLife(JOHN)\)
    (from 5 and 12)
 
  \end{enumerate}
  \end{enumerate}
\end{enumerate}
\end{document}
